---
title: "R Notebook DATA661Homework5"
output:
  html_notebook: default
  html_document: default
---

*Libraries*

```{r, warning=FALSE, echo=TRUE, message=FALSE}
library(ggplot2)
library(data.table)
library(dplyr)
library(ggrepel)
```



*Data Files & Table*

```{r,warning=FALSE, echo=TRUE, message=FALSE }
dat <- read.csv("https://raw.githubusercontent.com/Misterresearch/DATA661/master/EconomistData.csv")
head(dat)
```

*Exercise 1.1 -1.2*

Initial creation of plot, and changing points from the default black to blue.

```{r, warning=FALSE, echo=TRUE, message=FALSE}
ggplot(dat, aes(x = CPI, y = HDI)) + geom_point(color="blue")
```


*Exercise 1.3*

Setting colors based on region.

```{r}

ggplot(dat, aes(x = CPI, y = HDI, color = Region)) + geom_point()
```


*Exercise 1.4*

enlarging all points to size 2

```{r, warning=FALSE, echo=TRUE, message=FALSE}

ggplot(dat, aes(x = CPI, y = HDI, size = 2, color = Region)) + geom_point()
```


*Exercise 1.5*

Scaling size of points based on Rank

```{r, warning=FALSE, echo=TRUE, message=FALSE}
ggplot(dat, aes(x = CPI, y = HDI, size = HDI.Rank, color = Region)) + geom_point()
```


*Exercise 2.1-2.2*

Starting from initial plot again, then adding smoothing line 

```{r, warning=FALSE, echo=TRUE, message=FALSE}
ggplot(dat, aes(x = CPI, y = HDI, size = HDI.Rank)) + geom_point(color="blue")+ geom_smooth()
```

*Exercise 2.3*

Add regression line through plot using "stat_method"

```{r, warning=FALSE, echo=TRUE, message=FALSE}

ggplot(dat, aes(x = CPI, y = HDI, size = HDI.Rank)) + geom_point(color="blue")+ 
stat_smooth(method = lm)
```


*Exercise 2.4*

Now using geom_line, based on simple regression explicitly set in the code. 

```{r, warning=FALSE, echo=TRUE, message=FALSE}
predictHDI<- predict(lm(HDI ~ CPI, data = dat))

ggplot(dat, aes(x = CPI, y = HDI, size = HDI.Rank)) + geom_point(color="blue")+ 
geom_line(aes(y = predictHDI))
```

DID NOT ANSWER BONUS QUESTION

*Exercise 3.1*

Recreatio of initial plot with color set to region.

```{r, warning=FALSE, echo=TRUE, message=FALSE}
ggplot(dat, aes(x = CPI, y = HDI, color = Region)) + geom_point()
```

*Exercise 3.2*

Manual override of x label from variable name to custom text :Human Development Index"

```{r, warning=FALSE, echo=TRUE, message=FALSE}
ggplot(dat, aes(x = CPI, y = HDI, color = Region)) + geom_point() + scale_y_discrete(name="Human Development Index")
```


*Exercise 3.3*

Manual override of color=Region with a vector of custom variables. In order to work, the number of items in a vector must match the number of distinct values in a corresponding variable. 

```{r, warning=FALSE, echo=TRUE, message=FALSE}
ggplot(dat, aes(x = CPI, y = HDI, color=Region)) + geom_point() + scale_color_manual(values=c("black", "red", "green", "white", "orange", "purple"))
```
When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).


